---
title: "week 12"
author: "Alyssa Goldberg"
date: "June 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(fivethirtyeight)
library(r2d3)
```

```{r Get_Data}
gt<-google_trends
mchur<-mediacloud_hurricanes
mcstates<-mediacloud_states
mct<-mediacloud_trump
mco<-mediacloud_online_news
tvhur<-tv_hurricanes
tvnet<-tv_hurricanes_by_network
tvstates<-tv_states

#create tibble of storms for landmarking. Date is first landfall, category is category at landfall
storms<-tibble(date=c("2017-08-26", "2017-09-6","2017-09-10", "2017-09-09", "2017-09-20"), storm=c("harvey", "irma","irma", "jose", "maria"), category=c(4,5,5,5,5))
```
The 2017 Atlantic hurricane season was both active and severe.  Three storms struck US territory in quick succession: 

* Harvey  
* Irma  
* Jose  
* Maria  

Fivethirtyeight.com examined how these storms were covered by broadcast/cable news sources and how they trended on Google from the period: `r range(c(gt$date, mchur$date, mcstates$date, mct$date, tvhur$date, tvnet$date, tvstates$date))[1]` to `r range(c(gt$date, mchur$date, mcstates$date, mct$date, tvhur$date, tvnet$date, tvstates$date))[2]`

#Get Data  
Data comes from [fivethirtyeight.com's github](https://github.com/rudeboybert/fivethirtyeight)


# Explore Data
When did each of the storms first get mentioned in the media
```{r echo=TRUE}
#enlongen the tables for tidying and ggplot
tvhur<-tvhur %>% gather(key=storm, value=value, 2:5) 
mchur<-mchur %>% gather(key=storm, value=value, 2:5)

tvrange<-tvhur %>% 
  filter(value!=0) %>% 
  group_by(storm) %>% 
  filter(date==min(date)|date==max(date))

mcrange<-mchur %>% 
  filter(value!=0) %>% 
  group_by(storm) %>% 
  filter(date==min(date)|date==max(date))


```

```{r}
p<-ggplot(tvhur, aes(x=date, y=value, color=storm))
vline<-storms$date

p+geom_line()+
  geom_point(data=storms, aes(x=as.Date(date), y=1, color=storm))+
  geom_segment(data=storms, aes(x=as.Date(date), xend=as.Date(date), y=0, yend=1, color=storm), linetype="dashed")+
  theme_bw()

p1<-ggplot(mchur, aes(x=date, y=value, color=storm))

p1+geom_line()+
  geom_point(data=storms, aes(x=as.Date(date), y=max(mchur$value), color=storm))+
  geom_segment(data=storms, aes(x=as.Date(date), xend=as.Date(date), y=0, yend=max(mchur$value), color=storm), linetype="dashed")

```
```{r}
library(tidyverse)
library(fivethirtyeight)

#create tibble of storms for landmarking. Date is first landfall, category is category at landfall
#data from Google Search on each.  Irma has two dates, first landfall in Carribean and first landfall in Continental US.
storms<-tibble(date=c("2017-08-26", "2017-09-6","2017-09-10", "2017-09-09", "2017-09-20"), storm=c("harvey", "irma","irma", "jose", "maria"), category=c(4,5,5,5,5))



df=tv_hurricanes_by_network
df$query=as.character(df$query)

df<-df %>% gather(key=network, value=value, 3:6) %>% 
  group_by(date, network, query) %>% 
  summarise(value=mean(value, na.rm=TRUE)) %>% 
  ungroup() %>% 
    rename(storm=query) %>% 
  mutate(storm=tolower(gsub("Hurricane ", "", storm))) %>% 
  filter(value!=0) %>% 
  arrange(date)

range=range(df$value)

 p<-ggplot()+
   labs(title="Coverage by Network, Date and Storm",subtitle="The percent of sentences on a TV news channel on the\n given date that mention the hurricane in question\nblack dot = landfall date", x="", y=paste0("from ", range[1], "% to ", range[2],"%"), caption="data: fivethirtyeight\nviz: Alyssa Goldberg @WireMonkey\npart of the #TidyTuesday project by @thomas_mock")+
    geom_point(data=fortify(storms), aes(x=as.Date(date), y=1))+
    geom_line(data=df, aes(x=date, y=value, color=network))+
    coord_polar(theta="x", clip="off")+
    theme_bw()+
    facet_wrap(~storm)
 
 png(filename = "tt12.png",type = "cairo", units="in", res=100, width = 5,height = 5)
 p
dev.off()

```

